---
title: "Single Effect Regression Benchmarks"
format:
  html: 
    code-fold: true
---

```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = '~/Research/logistic-susie-benchmarks/')
```

# Single Effect Regression Benchmarks

## Introduction

There are many ways to approximate the logistic SER. Here we document the relative performance of these different approximations. We investigate a range of simulation scenarios, varying the properties of the design matrix (correlation structure, sparsity), the number of true causal effect ($L = 1, 3, 5$), the strength of effect, etc.

```{r}
library(targets)
library(tidyverse)

cs_scores <- tar_read(ser_cs_coverage, store = '_targets/')
fit_methods <- unique(cs_scores$fit_method)

plot_colors <- c(
  'glm_ser' = 'aquamarine1', 
  'glm_ser2' = 'aquamarine3', 
  'uvb_ser' = 'brown1', 
  'uvb_ser2' = 'brown3', 
  'vb_ser' = 'darkgoldenrod1', 
  'vb_ser2' = 'darkgoldenrod3',
  'vb_ser_corrected' = 'darkorange1',
  'veb_ser' = 'darkorchid1',
  'quad_ser' = 'deeppink1',
  'uvb_ser_re' = 'palevioletred3',
  'jj_abf_ser' = 'royalblue4'
)
plot_colors <- plot_colors[names(plot_colors) %in% fit_methods]
method_order = names(plot_colors)

# coverage for each L
plot_data <- cs_scores %>%
  ungroup() %>%
  group_by(fit_method, L) %>%
  # distribution of mean(covered) over 20 bootstrap iterations
  mutate(coverage = list(purrr::map_dbl(1:20, ~mean(sample(covered, replace = T))))) %>%
  select(fit_method, coverage, L) %>%
  unnest_longer(coverage)

plot_data %>% dplyr::pull(fit_method) %>% unique()
```

```{r}
plot <- plot_data %>%
  ggplot(aes(x=factor(fit_method, level = names(plot_colors)), y=coverage, fill=fit_method)) +
  geom_boxplot(outlier.shape=NA) + facet_wrap(vars(L)) +
  scale_fill_manual(values=plot_colors) +
  theme(axis.text.x = element_blank())
plot
```



```{r}
# coverage conditional on CS size
plot_data <- cs_scores %>%
  ungroup() %>%
  hoist(cs, size='size') %>%
  mutate(size_bin = cut(size, c(0, 1, 5, 10, 50))) %>%
  group_by(fit_method, size_bin, L) %>%
  mutate(coverage = list(purrr::map_dbl(1:50, ~mean(sample(covered, replace = T))))) %>%
  select(fit_method, size_bin, coverage, L) %>%
  unnest_longer(coverage)

plot <- plot_data %>% 
  ggplot(aes(x=size_bin, y=coverage, fill=factor(fit_method, level=method_order))) +
  geom_boxplot(outlier.shape=NA) + facet_grid(rows=vars(L)) +
  scale_fill_manual(values=plot_colors)

plot
```


```{r}
roc_plot <- tar_read(ser_pip_roc, store='_targets/')
roc_plot + xlim(0, 0.2) + ylim(0, 0.4) + scale_color_manual(values=plot_colors)
```

```{r}
# coverage of SERs for different true L
cs_scores %>%
  ungroup() %>%
  hoist(cs, size='size') %>%
  group_by(fit_method, L) %>%
  mutate(coverage = list(purrr::map_dbl(1:100, ~mean(sample(covered, replace = T))))) %>%
  select(fit_method, L, coverage) %>%
  unnest_longer(coverage) %>%
  ggplot(aes(x=factor(L), y=coverage, fill=factor(fit_method, levels = method_order))) +
  scale_fill_manual(values=plot_colors) + 
  geom_boxplot(outlier.shape=NA)
```

